A Decision Procedure for (Co)datatypes in SMT Solvers

### 2014-12-22 tex/codatadp.tex
minor

\begin{figure}[b!]
\normalsize
\centering
\begin{tabular}{c}
\(
\inferrule{
  \delta \in \Data
  \quad
  t : \delta
  \quad
  %\ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
  \Val \ec{t} = \MU x.\; u
  \quad
  x \in \FV( u )
}{
  \bot
}
\)
\rn{Acyclic}
\\[4\jot]
\(
\inferrule{
 \delta \in \Codata
 \quad
 t, u : \delta
 \quad
 %\tpath{\Ec}{\ec{t}} = \tpath{\Ec}{\ec{u}}\sigma \neq \emptyset
 \Val \ec{t} \aequiv \Val \ec{u}
}{
 \Ec := \Ec,\: t \teq u
}
\)
\rn{Unique}
\end{tabular}
\caption{\,Derivation rules for acyclicity and uniqueness%.
}
\label{fig:ab-rules}
\end{figure}

@@

\begin{figure}[tbh]
\normalsize
\centering
\begin{tabular}{c}
\(
\inferrule{
  t : \delta \in \tEc
  \quad
  ( \const s( t ) \in \tEc \text{ and } \const s \in \Sel^\delta )
  \text{ or }
  ( \delta \in \Data \text{ and } \delta \text{ finite} )
}{
  \Ec := \Ec,\: t \teq \const{C}_1(\const s^1_1( t ), \ldots,\const s^{n_1}_1( t ) ) \ror \cdots \ror \Ec := \Ec,\: t \teq \const{C}_m(\const s^1_m( t ), \ldots,\const s^{n_m}_m( t ) )
}
\)
\rn{Split}
\\[5\jot]
\(
\inferrule{
  \delta \in \Codata
  \quad
  \left|\delta\right| = 1
  \quad
  t, u : \delta \in \tEc
}{
  \Ec := \Ec,\: t \teq u
}
\)
\rn{Singleton}
\end{tabular}
\caption{\,Derivation rules for branching%.
%All pairs of terms whose type has cardinality 1 are entailed to be equal (the degenerate case).
%Constructors must be assigned for all terms $t$ if has a selector is applied to it, or if $t$ has finite datatype type.
}
\label{fig:split-rule}
\end{figure}

\paragraph{Phase 3: Branching.}
If a selector is applied to a term $t$, or if $t$'s type is a finite datatype,
the equivalence class of $t$ must contain an application of one of
$\delta$'s constructors
$\Ctr^\delta = \{ \const{C}_1, \ldots, \const{C}_m \}$.
This is enforced in the third phase by the \rn{Split} rule.
A companion rule, \rn{Singleton}, handles the degenerate case where two terms
from $\tEc$ have a corecursive singleton type, in which case they must be
equivalent. Both the finiteness assumption on the datatype in \rn{Split} and
the cardinality-1 constraint on the codatatype in \rn{Singleton} are 
are computed statically
based on a recursive computation of the cardinalities of the
constructors' argument types.

>>>>

\begin{figure}[b!]
\normalsize
\centering
\begin{tabular}{c}
\(
\inferrule{
  t : \delta
  \quad
  \delta \in \Data
  \quad
  %\ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
  \Val \ec{t} = \MU x.\; u
  \quad
  x \in \FV( u )
}{
  \bot
}
\)
\rn{Acyclic}
\\[4\jot]
\(
\inferrule{
 t, u : \delta
 \quad
 \delta \in \Codata
 \quad
 %\tpath{\Ec}{\ec{t}} = \tpath{\Ec}{\ec{u}}\sigma \neq \emptyset
 \Val \ec{t} \aequiv \Val \ec{u}
}{
 \Ec := \Ec,\: t \teq u
}
\)
\rn{Unique}
\end{tabular}
\caption{\,Derivation rules for acyclicity and uniqueness%.
}
\label{fig:ab-rules}
\end{figure}

@@

\begin{figure}[tbh]
\normalsize
\centering
\begin{tabular}{c}
\(
\inferrule{
  t : \delta \in \tEc
  \quad
  \bigl( \const s( t ) \in \tEc \text{ and } \const s \in \Sel^\delta \bigr)
  \text{ or }
  \bigl( \delta \in \Data \text{ and } \delta \text{ finite} \bigr)
}{
  \Ec := \Ec,\: t \teq \const{C}_1(\const s^1_1( t ), \ldots,\const s^{n_1}_1( t ) ) \ror \cdots \ror \Ec := \Ec,\: t \teq \const{C}_m(\const s^1_m( t ), \ldots,\const s^{n_m}_m( t ) )
}
\)
\rn{Split}
\\[5\jot]
\(
\inferrule{
  t, u : \delta \in \tEc
  \quad
  \delta \in \Codata
  \quad
  \left|\delta\right| = 1
}{
  \Ec := \Ec,\: t \teq u
}
\)
\rn{Singleton}
\end{tabular}
\caption{\,Derivation rules for branching%.
%All pairs of terms whose type has cardinality 1 are entailed to be equal (the degenerate case).
%Constructors must be assigned for all terms $t$ if has a selector is applied to it, or if $t$ has finite datatype type.
}
\label{fig:split-rule}
\end{figure}

\paragraph{Phase 3: Branching.}
If a selector is applied to a term $t$, or if $t$'s type is a finite datatype,
the equivalence class of $t$ must contain an application of one of
$\delta$'s constructors
$\Ctr^\delta = \{ \const{C}_1, \ldots, \const{C}_m \}$.
This is enforced in the third phase by the \rn{Split} rule.
A companion rule, \rn{Singleton}, handles the degenerate case where two terms
from $\tEc$ have a corecursive singleton type
(Section~\ref{sec:the-theory-of-co-datatypes}), in which case they must be
equivalent. Both the finiteness assumption on the datatype in \rn{Split} and
the cardinality-1 constraint on the codatatype in \rn{Singleton} are 
are computed statically
based on a recursive computation of the cardinalities of the
constructors' argument types.
